# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Loading and preprocessing data:
```{r loading,echo=TRUE}
data<-read.csv('activity.csv')
data$date<-as.Date(data$date, format="%Y-%m-%d")

```


## What is mean total number of steps taken per day?
Creating the histogram of the average total number of steps taken each day:
```{r steps,echo=TRUE}
datasum<-tapply(data$steps, data$date,sum,na.rm=TRUE)
barplot(datasum,xlab="Date",ylab="Total number of steps",main="Total number of steps taken each day")
```

The mean of the average total number of steps taken per day is `r mean(datasum,na.rm=TRUE)`, while the median is `r median(datasum,na.rm=TRUE)`.

## What is the average daily activity pattern?
Creating a plot of the average steps taken in each 5-minute interval.
```{r daily,echo=TRUE}
datasteps<-tapply(data$steps,data$interval,mean,na.rm=TRUE)
bla<-cbind(interval=unique(data$interval),steps=datasteps)
bla2<-strptime(unique(sapply(data$interval, formatC, width = 4, flag = 0)),format="%H%M")
plot(bla2,bla[,2],type="l",xlab="Hour of the day",ylab="Average steps",main="Daily activity")
```

The interval containing the maximum number of steps is the one starting at `r substr(bla2[which.max(datasteps)],12,19)`


## Inputing missing values
Inserting the mean number of steps for the interval, when the data is missing. There are `r sum(is.na(data))` rows with missing data.
```{r missing,echo=TRUE}
data2<-data
for (i in 1:nrow(data2)){
  if(is.na(data2$steps[i])){
    x<-bla[bla[,1]==data2$interval[i]]
    data2$steps[i]<-x[2]
  }
}
head(data2)
```

Creating a histogram with the filled-out dataset.
```{r newhist,echo=TRUE}
datasum2<-tapply(data2$steps, data2$date,sum,na.rm=TRUE)
barplot(datasum2,xlab="Date",ylab="Total number of steps",main="Total number of steps taken each day")
```

This histogram is not different than the previous version.

The mean of the average total number of steps taken per day is `r mean(datasum2,na.rm=TRUE)`, while the median is `r median(datasum2,na.rm=TRUE)`. These are also the same as those of the original dataset.


## Are there differences in activity patterns between weekdays and weekends?

Creating a new factor variable with two levels.

```{r weekends,echo=TRUE}
days<-c()
for (i in 1:nrow(data2)){
if (weekdays(as.Date(data2$date[i]))=="Saturday" || weekdays(as.Date(data2$date[i]))=="Sunday"){
  days[i]<-"weekend"
}
else{days[i]<-"weekday"}

}
data2$day<-factor(days)
```



Time series plot for weekdays and weekends.
```{r plot,echo=TRUE}
library(data.table)
data2<-data.table(data2)
data3<-data2[,mean(steps),by="interval,day"]
library(lattice)
xyplot(data3$V1~data3$interval| data3$day,type="l",layout=c(1,2),xlab="Interval",ylab="Number of steps")

```